---
title: "Data605_Assignment04"
author: "Mahmud Hasan Al Raji"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This assignment is done to build and visualize eigenimagery that accounts for 80% of the variability of the given image data set. To do so, following steps are being followed:

(1) Uploading the image files as a list that containing the image features i.e. names,pixels etc.
(2) Converting the multi-dimensional data list into an array 
(3) Transforming the array into a column vector. The vector is formed by splitting each image into R,G,B components and concatenate these components into that vector.
(4) Creating an empty matrix and then loading RGB vectors into the matrix and then creating a data frame with the transpose of the data matrix.
(5) Normalizing the data set using the mean value (=0) and standard deviation (=1) and then the correlation matrix is computed.
(6) Calculating the eigenvectors and eigenvalues of the covariance matrix and then determining the principal components of the images.
(7) Finally, identifying the principal components that accounted for 80% of the variability in the data set and computing the new data set by projecting the original data into the newly found eigenspace.

# Required Libraries

```{r message=FALSE}
library(jpeg)
library(OpenImageR)

```

# Creating list for images' features (pixels included)

```{r }
files<-list.files("C:\\Users\\Acer\\Documents\\jpg",pattern="\\.jpg",full.names = TRUE)
```

# Loading data into an array

```{r}
for (file in files) {
  img <- readImage(file)  
}

```

# Vectorizing with RGB components of the images 

```{r}
flat <- matrix(0, length(files), prod(dim(img))) 
for (i in 1:length(files)) {
  im <- readJPEG(files[i])
  r  <- as.vector(im[,,1])
  g  <- as.vector(im[,,2])
  b  <- as.vector(im[,,3])
  
  flat[i,] <- t(c(r, g, b))
}

shoes=as.data.frame(t(flat))
```

# Normalizing the dataset using the mean value (=0) and standard deviation (=1):

```{r }
scaled<-scale(shoes, center = TRUE, scale = TRUE)
mean.shoe<-attr(scaled, "scaled:center") #saving for classification
std.shoe<-attr(scaled, "scaled:scale")  #saving for classification...later
```

# Calculating Covariance (Correlation)

Finding correlation matrix showing the correlations between pairs of variables. This correlation matrix will be used to get the eigencomponents.
```{r}
Sigma_<-cor(scaled)
```

# Getting the Eigencomponents:Calculating the eigenvalues of the covariance matrix

```{r}
eig          <- eigen(Sigma_)
eigenvalues  <- eig$values
eigenvectors <- eig$vectors

```

# Choosing the number of principal components

```{r}
prop.var <- eigenvalues / sum(eigenvalues)
cum.var  <- cumsum(eigenvalues) / sum(eigenvalues)
thres    <- min(which(cum.var > .80))
thres
```

# Eigenshoes accounted for 80% of the variability

```{r}
scaling<-diag(eigenvalues[1:thres]^(-1/2)) / (sqrt(nrow(scaled)-1))
eigenshoes<-scaled%*%eigenvectors[,1:thres]%*%scaling
imageShow(array(eigenshoes[,1], dim(img)))
```

# Second Eigenshoe 
```{r}
imageShow(array(eigenshoes[,2], dim(img)))
```

# visualization the proportion of variance explained by each of the components by means of a scree plot 

```{r}
mypca <- prcomp(shoes, center = TRUE, scale= TRUE)
pca.var <- mypca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100,1)
barplot(pca.var.per, main = "Scree Plot", xlab = "Principal Component", ylab = "Percent Variation")
```

# Computing new dataset

```{r}
data_new<-as.data.frame(t(t(eigenshoes)%*%scaled))
```

It is found that there are 3 principal components among all the  components are accounted for 80% of the variability. 

References: 

(1) https://rpubs.com/dherrero12/543854
(2) https://rpubs.com/R-Minator/eigenshoes




